---
title: "Compare Effect Sizes"
author: "Agaz Wani"
date: "2/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(hrbrthemes)
```


### Code to compare the effect sizes of the discovery and independent cohort. Here methylation is a function of PTSD

```{r}
disc <- read.csv("../../../Combined Data/2021-11-27_19-16-35/lm_results_testset_ptsd.csv")
message("dimesion of discovery :")
dim(disc)

indp <- read.csv("../../Kurt output/VA_MRS/AdditionalResult1/effect_sizes_ptsd_whenEWASPC.csv")
colnames(indp) <- paste0(colnames(indp), "_indp")
message("dimesion of independent :")
dim(indp)
```

### Merge data
```{r}
es_comb <- merge(disc, indp,
                 by = 1)
# head(es_comb)
# View(es_comb)
```


### Count positive and negative directions in discovery and independent cohorts 
```{r}
cbind( es_comb %>% count(Estimate < 0),
       es_comb %>% count(Estimate_indp < 0))
```
### Pull out features with same and different directions 
```{r}

fea_diff_dir <- es_comb %>% 
  filter(sign(Estimate) != sign(Estimate_indp))
dim(fea_diff_dir)
# View(fea_diff_dir)

fea_same_dir <- es_comb %>% 
  filter(sign(Estimate) == sign(Estimate_indp))
dim(fea_same_dir)
# View(fea_diff_dir)

```



```{r}
# Horizontal version
lp_th <- theme(plot.title = element_text(size = 20, hjust = 0.5),
              axis.title = element_text(size =18),
              axis.text = element_text(size = 10, face = 'bold'),
              legend.title = element_text( size = 16),
              legend.text = element_text(size = 14),
              panel.grid.major.y = element_blank(),
              panel.border = element_blank(),
              axis.ticks.y = element_blank()
  )


line_plot <- function(df, x, y, y1, xlab="", ylab="", title="", legend = TRUE){
  g <- ggplot(df) +
    geom_segment( aes_string(x=x, xend=x, y=y, yend=y1), color="grey") +
    geom_point( aes_string(x=x, y=y, color = "y"), size=4 ) +
    geom_point( aes_string(x=x, y=y1, color = "y1"), size=4 ) +
    scale_color_manual(values = c("orange", "deepskyblue4"),
                       guide  = guide_legend(), 
                       name   = "Group", 
                       labels = c("Discovery", "Independent")) +
    theme_light()
  
  if(legend){
    g <- g + theme(legend.position = "top",
          panel.border    = element_blank())
  }else{
   g <- g +  theme(
    legend.position = "none",
  )
  }
   g + coord_flip() +
    ggtitle(title) +
    lp_th +
    xlab(paste0(xlab, " (", length(df[[x]]), ")" )) +
    ylab(ylab)
}
```


```{r, fig.width=6, fig.height=13}
fea_diff_dir <- fea_diff_dir %>% 
   mutate(esmean = mean(c(Estimate, Estimate_indp))) %>%
  arrange(esmean)

n = ceiling(nrow(fea_diff_dir) / 2)
nr <- nrow(fea_diff_dir)

df_ls <- split(fea_diff_dir, rep(1:ceiling(nr/n), each=n, length.out=nr))
lapply(df_ls, dim)

leg <- c(TRUE, FALSE)
  
p_ls <- lapply(seq_along(df_ls), function(i){
    line_plot(df = df_ls[[i]],  x = "Feature", y= "Estimate",
          y1 =  "Estimate_indp", xlab = "Features",
          ylab = 'Beta Estimate', legend = leg[i],
          title = "Feature with different direction")
})

# p_ls

```


```{r, fig.height=14, fig.width=12}
cowplot::plot_grid(p_ls[[1]], p_ls[[2]])
```
### Features with same direction of effect
```{r, fig.height=13, fig.width=7}
line_plot(df = fea_same_dir,  x = "Feature", y= "Estimate",
          y1 =  "Estimate_indp", xlab = "Features",
          ylab = 'Beta Estimate', legend = TRUE,
          title = "Feature with same direction")
```


### Now load methylation vs age and smoking results. Here is methylation as a function of age and smoking scores (gender and ancestry mPCs) as covariates 
```{r}
disc_age_sm <- local(mget(load("G:/PGC ML/Combined Data/2021-11-27_19-16-35/age_smoking_effect_sizes.Rdata")))
# names(disc_age_sm)
lapply(disc_age_sm, dim)

# lapply(disc_age_sm, head)
```

### Independent cohort
```{r}
indp_age_sm <- local(mget(load("G:/PGC ML/RCode/Kurt output/VA_MRS/AdditionalResult1/age_smoking_effect_sizes_whenEWASPC.Rdata")))
# names(indp_age_sm)
lapply(indp_age_sm, dim)

# add to col names
indp_age_sm <- lapply(indp_age_sm, function(x){
  colnames(x) <- paste0(colnames(x), "_indp")
  x
})

# lapply(indp_age_sm, head)

```

### Now merge effect sizes of age and smoking of discovery and independent cohort
```{r}

comb_age_sm <- Map(function(x, y) merge(x, y, by.x = 1, by.y = 0), disc_age_sm, indp_age_sm)
lapply(comb_age_sm, dim)
# lapply(comb_age_sm, head)
# lapply(comb_age_sm, names)
```

### How many have positive and negative direction for age and smoking
```{r}
lapply(names(comb_age_sm), function(x){
  print(x)
  cbind(comb_age_sm[[x]] %>% count(Estimate < 0), 
        comb_age_sm[[x]] %>% count(Estimate_indp < 0))
}) 
```



### Lets get the cpgs with different direction for age and smoking
```{r}
cpgs_wd_dff <- lapply(comb_age_sm, function(x){
  x %>% 
  filter(sign(Estimate) != sign(Estimate_indp))
})

lapply(cpgs_wd_dff, dim)
```

### Plot effect size of cpgs for age and smoking
```{r}
leg1 <- c(TRUE, TRUE, FALSE, FALSE)
titles <- c("Age", "Age with covariates", "Smoking score", "Smoking score with covariates" )

p_ls_age_sm <- lapply(seq_along(cpgs_wd_dff), function(i){
    line_plot(df = cpgs_wd_dff[[i]],  x = "Feature", y= "Estimate",
          y1 =  "Estimate_indp", xlab = "Features",
          ylab = 'Beta Estimate', legend = leg1[i],
          title = titles[i] )
})
names(p_ls_age_sm) <- names(cpgs_wd_dff)

```


```{r, fig.width=12, fig.height=10}
cowplot::plot_grid(p_ls_age_sm$es_age_covar, p_ls_age_sm$es_smos_covar)

```

### Are there any common cpgs with different direction with covariates 
```{r}
message("Common in age and smoking with different directions:")
table(cpgs_wd_dff$es_age_covar$Feature %in% cpgs_wd_dff$es_smos_covar$Feature)
```

### Are there any common cpgs with different direction *without* covariates 
```{r}
message("Common in age and smoking with different directions:")
table(cpgs_wd_dff$es_age$Feature %in% cpgs_wd_dff$es_smos$Feature)
```


### Looking at mean methylation for age and smoking
```{r}

mean_meth_disc <- read.table("G:/PGC ML/RCode/PGCML/data/Linear model age and smoking summary.txt",
                             fill = T)

mean_meth_indp <- read.table("G:/PGC ML/RCode/Kurt output/VA_MRS/AdditionalResult1/Linear model age and smoking summary_whenEWASPC.txt", fill = T)


```

### get age and smoking rows only
```{r}
ls_age_sm <- lapply(list(mean_meth_disc,mean_meth_indp), function(x){
  x %>% filter(grepl("Age|Smo", V1))
})  
                           
new_n <- c("Age", "Age_wd_Covars", "SmoS", "SmoS_wd_Covars")
ls_age_sm <- lapply(ls_age_sm, function(x){
  x$V1 <- new_n
  x <- x[, 1:5]
  colnames(x) <- c("Feature", 'Estimate', 'Std. Error',  't value', 'Pr(>|t|)')
  # x$Estimate <- round(x$Estimate, 4)
  x
})

names(ls_age_sm) <- c("Disc", "Indp")

# add to the names of indp
names(ls_age_sm$Indp) <- c("Feature", paste0(colnames(ls_age_sm$Indp[, -1]), "_indp"))

# merge
comb_meanm_st <- merge(ls_age_sm$Disc, ls_age_sm$Indp, by = 1)
print(comb_meanm_st[, c(1,2,5,6,9)])
```


```{r, fig.height=5, fig.width=8}
line_plot(df = comb_meanm_st,  x = "Feature", y= "Estimate",
          y1 =  "Estimate_indp", xlab = "Features",
          ylab = 'Beta Estimate', legend = TRUE,
          title = "Effect size of age and smoking")
```

