---
title: "Downstream analysis"
author: "Agaz Wani"
date: "10/25/2021"
output: html_document
---

```{r setup, include=FALSE}
 # ⌘+⌥+I / Ctrl+Alt+I splits a code chunk.
knitr::opts_chunk$set(echo = TRUE)
```


### Code to do enrichment analysis

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(missMethyl)
library(ggplot2)
library(ggsignif)
library(dplyr)
library(tidyr)
library(corrplot)
library(cowplot)
```


### risk scores on test set
```{r}
# 2021-10-02_23-12-43 previous path
adj_df <- read.csv("G:/PGC ML/Combined Data/2021-11-27_19-16-35//risk scores ptsdpm test data.csv")

adj_df <- adj_df %>% dplyr::rename(CurrentPTSD = ptsdpm, 
                            TraumaNumber = Traumanum,
                            ChildhoodMaltreatment = Childhood_Mt)
adj_df$CurrentPTSD <- as.factor(adj_df$CurrentPTSD)
```


### To scale data remove the risk scores and ptsd
```{r}
test_sc <- adj_df[, -which(names(adj_df) %in% (c("risk_score", "CurrentPTSD")))]
dim(test_sc)

```


### Lifetime 
```{r}
lifetime <- read.csv("G:/PGC ML/Combined Data/2021-10-02_23-12-43/Important_features_ptsdlife_adjusted.csv")
```


### function to scale the data betweem 0-1
```{r}
maxmin <- function(x, na.rm=TRUE){
  if(is.vector(x)==TRUE){
    maxs <- max(x, na.rm = na.rm)
    mins <- min(x, na.rm = na.rm)
    scale(x,center=mins,scale=maxs-mins)
  } else {
    maxs <- apply(x, 2,max)
    mins <- apply(x, 2,min)
    scale(x, center = mins, scale = maxs - mins)
  }
}


```


### We will scale trauma and childhoodmaltreatment columns
```{r}
scale_cols <- c("TraumaNumber", "ChildhoodMaltreatment")
scale_df <- test_sc[, which(names(test_sc) %in% (scale_cols))]
scale_df <- data.frame(maxmin(x = scale_df, na.rm = T))

other_df <- test_sc[, -which(names(test_sc) %in% (scale_cols))]

adj_sc_df <- cbind(other_df, scale_df)

```


### Now calculate risk scores
```{r}
weights <- read.csv("G:/PGC ML/Combined Data/2021-11-27_19-16-35/Important_features.csv")

weights$Feature <- ifelse(weights$Feature == 'Traumanum', 'TraumaNumber',
                          ifelse(weights$Feature == 'Childhood_Mt',
                                 'ChildhoodMaltreatment', weights$Feature))


table(colnames(adj_sc_df) == weights$Feature)

risk_score <- apply(adj_sc_df, 1, function(x) sum(x * weights$Importance))

adj_sc_df$CurrentPTSD <- adj_df$CurrentPTSD

```


### Get beta-estimates
```{r}
b_estimates = list()

for (i in 1:300){
 x <- colnames(adj_sc_df)[i]
 f <- paste0(x, '~', 'CurrentPTSD')
 es <- lm(f, data = adj_sc_df)
 summary(es)
 b_estimates[[length(b_estimates) +1 ]] <- es$coefficients[2]
}

names(b_estimates) <- weights$Feature

# add beta-estimates to the df
weights$beta_est <- unlist(unname(b_estimates))

```

### Add the calculated risk scores to df
```{r}
adj_sc_df$risk_score_imp_beta <-  apply(adj_sc_df[, -301], 1, function(x) sum(x * weights$Importance*weights$beta_est))

adj_sc_df$risk_score <- risk_score

```





### calculate mean methylation of samples for significant cpgs
```{r}
p <- c("^cg", "^ch")

adj_df <- adj_df %>%
  mutate(mean_meth = rowMeans(select(., starts_with(c('cg','ch.')))))

adj_sc_df <- adj_sc_df %>%
  mutate(mean_meth = rowMeans(select(., starts_with(c('cg','ch.')))))


```



### Correlation plot
```{r, fig.width=12, fig.height=6}
cols <- c("Gender", "TraumaNumber", "ChildhoodMaltreatment",
          "risk_score_imp_beta", "risk_score", "CurrentPTSD", 
          "mean_meth")

new_names <- c("Gender", "Cumulative trauma", "Childhood maltreatment", 
              "Risk score ES",  "Risk score", "Current PTSD", 
              "Mean methylation")

# # train
# # adj_sc_df <- adj_sc_df %>% select(cols)
# # adj_sc_df$ptsdpm <- as.numeric(adj_sc_df$ptsdpm)
# # colnames(adj_sc_df) <- new_names
# # cor_vec <- cor(adj_sc_df)
# # 
# # trainRes = cor.mtest(df_cor, conf.level = 0.95)
# # 
# # matrix(p.adjust(as.vector(as.matrix(trainRes$p)), method='fdr'),
# #        ncol=6)
# # 
# # # test ------------------
# # df_cor <- adj_df %>% select(cols)
# # df_cor$ptsdpm <- as.numeric(df_cor$ptsdpm)
# # 
# # colnames(df_cor) <- new_names
# # 
# # cor_vec <- cor(df_cor)
# # cor_vec
# 
# colrs <- colorRampPalette(c("#0072B5FF", '#7F0000',
#            '#FF7F00', 'orange',
#            'cyan', '#007FFF', 'blue', '#00007F',
#            "#BC3C29FF"))
# #pvalue
# testRes = cor.mtest(df_cor, conf.level = 0.95)
# 
# matrix(p.adjust(as.vector(as.matrix(testRes$p)), method='fdr'),
#        ncol=6, )
# 
# 
# p_train <- corrplot(cor_vec_train,p.mat = trainRes$p, diag = TRUE, method = "pie",
#               type = "upper",  number.cex = 0.8,
#               tl.col = "black", tl.srt = 45,
#               tl.cex = 1.7, cl.cex = 1.5,
#               sig.level = c(0.001, 0.01, 0.05), 
#               pch.cex = 1.5,insig = 'label_sig',
#               cl.ratio = 0.4,
#               col = colrs(10),
#               title = "Training data")
# 
# p_test <- corrplot(cor_vec,p.mat = testRes$p, diag = TRUE, method = "pie",
#               type = "upper",  number.cex = 0.8,
#               tl.col = "black", tl.srt = 45,
#               tl.cex = 1.7, cl.cex = 1.5,
#               sig.level = c(0.001, 0.01, 0.05), 
#               pch.cex = 1.5,insig = 'label_sig',
#               cl.ratio = 0.4,
#               col = colrs(10),
#               title = "Test data")
# 
# cor(adj_df$risk_score, as.numeric(adj_df$ptsdpm))
# 
# cowplot::plot_grid(plot(p_train), plot(p_test))

```

## Function to get correlation and adjusted pvalue. Input is data frame, columns we want to get correlation for and new names to assign for columns. 
```{r}
get_correlation <- function(df, cols, new_names){
  df <- df %>% select(cols)
  indx <- sapply(df, is.factor)
  df[indx] <- lapply(df[indx], function(x) as.numeric(as.character(x)))
  colnames(df) <- new_names
  cor_df <- cor(df)
  p = cor.mtest(df, conf.level = 0.95)

  # adjusted pvalues
  p_adj <- matrix(p.adjust(as.vector(as.matrix(p$p)),
                           method='fdr'),
                  ncol=ncol(cor_df))
  rownames(p_adj) <- new_names
  colnames(p_adj) <- new_names

  df_ls <- list(cor_df, p_adj)
  names(df_ls) <- c("cor", "padj")
  return(df_ls)
}

# color palette
colrs <- colorRampPalette(c("#0072B5FF", '#7F0000',
                            '#FF7F00', 'orange',
                            'cyan', '#007FFF', 'blue', '#00007F',
                            "#BC3C29FF"))
```


## Now call the function to get correlation for unadjusted data
```{r, fig.width=8, fig.height=5}
# covariate adjusted data
# cor_adj_df <- get_correlation(df = adj_df, cols = cols,
#                                 new_names = new_names)

# covariate adjusted and scaled (phenotypes)data
cor_scaled_df <- get_correlation(df = adj_sc_df, cols = cols,
                                new_names = new_names)



# p_adj <-  corrplot(cor_adj_df$cor,p.mat = cor_adj_df$padj,
#          diag = TRUE, method = "pie",
#          type = "upper",  number.cex = 0.8,
#          tl.col = "black", tl.srt = 45,
#          tl.cex = 1.5, cl.cex = 1.5,
#          sig.level = c(0.001, 0.01, 0.05),
#          pch.cex = 1.5,insig = 'label_sig',
#          cl.ratio = 0.4,
#          col = colrs(10),
#          title = "")

# grid.echo()
# p_unadj <- grid.grab()




p_adj_sc <- corrplot(cor_scaled_df$cor,p.mat = cor_scaled_df$padj,
                  diag = TRUE, method = "pie",
                  type = "upper",  number.cex = 0.8,
                  tl.col = "black", tl.srt = 45,
                  tl.cex = 1.5, cl.cex = 1.5,
                  sig.level = c(0.001, 0.01, 0.05),
                  pch.cex = 1.5,insig = 'label_sig',
                  cl.ratio = 0.4,
                  col = colrs(10),
                  title = "Covariate adjusted and scaled")

# grid.echo()
# p_adj <- grid.grab()

# plot_grid(p_unadj, NULL, p_adj, rel_widths = c(1,-0.1,1),
#           labels = c('A', '', 'B'), align = "hv", nrow = 1)

```

### Theme and linear model plot function

```{r}
# theme setting
th <-   theme(plot.title = element_text(size = 20, hjust = 0.5),
              axis.title = element_text(size =18),
              axis.text = element_text(size = 16),
              legend.title = element_text( size = 16),
              legend.text = element_text(size = 14))


# function to plot with lm line
lm_plot <- function(df, x, y, xlab, ylab, title, group='ptsdpm',
                    model = 'full', legend='', fig_legend = TRUE){
  if (model == 'full'){
    p <- ggplot(df,aes_string(x, y)) +
      geom_point( color="blue", size = 3) +
      geom_smooth(method='lm', linetype="dashed",
                  color="darkred", fill="blue", size = 1.5)
  }else{
    p <- ggplot(df,aes_string(x, y, color = group)) +
      geom_point( size = 3) 
    
    if(fig_legend){
      print("=================")
      p <- p + scale_color_manual(legend,
                         values = c("0" = "#0072B5FF", "1" = "#BC3C29FF"))+
        geom_smooth(aes(group = group, color=group),
                  method='lm', linetype="solid",
                  size = 1.5) #color="blue", fill="blue",
    }else{
      p <- p + scale_color_manual(legend,
                         values = c("0" = "#0072B5FF", "1" = "#BC3C29FF"))+
        theme(legend.position = 'none')+
        geom_smooth(aes(group = group, color=group),
                  method='lm', linetype="solid",
                  size = 1.5) #color="blue", fill="blue",
      
    }


  }

  p <- p + ggtitle(title)+
    xlab(xlab) +
    ylab(ylab) +
    theme_classic()+
    th
}


```

# linear model
```{r}

sink("G:/PGC ML/Combined Data/2021-10-02_23-12-43/Linear model PTSD summary.txt")

# current PTSD
lmr_adj <- lm(risk_score_imp_beta~CurrentPTSD, data = adj_sc_df) 
summary(lmr_adj)

lmr_scaled <- lm(risk_score~CurrentPTSD, data = adj_sc_df)
summary(lmr_scaled)


# lifetime PTSD
# lmr_unadj_l <- lm(risk_score~LifetimePTSD, data = unadj_df)
# summary(lmr_unadj_l)


# lmr_adj_l <- lm(risk_score~LifetimePTSD, data = adj_df)
# summary(lmr_adj_l)

# risk scores vs mean methylation
lmr_meth <- lm(mean_meth~risk_score_imp_beta, data = adj_sc_df)
summary(lmr_meth)

lmr_meth_sc <- lm(mean_meth~risk_score, data = adj_sc_df)
summary(lmr_meth_sc)


# cumulative trauma vs mean methylation
# lmr_ct_meth <- lm(mean_meth~TraumaNumber, data = adj_df)
# summary(lmr_ct_meth)

lmr_ct_meth_sc <- lm(mean_meth~TraumaNumber, data = adj_sc_df)
summary(lmr_ct_meth_sc)


# childhood maltreatment vs mean methylation
# lmr_chd_meth <- lm(mean_meth~ChildhoodMaltreatment, data = adj_df)
# summary(lmr_chd_meth)


lmr_chd_meth_sc <- lm(mean_meth~ChildhoodMaltreatment, data = adj_sc_df)
summary(lmr_chd_meth_sc)

sink()

```

### Separate cases and controls
```{r}
cases <- adj_df %>% filter(CurrentPTSD == 1)
message("cases:")
dim(cases)

controls <- adj_df %>% filter(CurrentPTSD == 0)
message("controls:") 
dim(controls)

```

# lm for mean methylation and risk scores
```{r, fig.width=11, fig.height=4}

p_meth_sep <- lm_plot(df = adj_sc_df, x = 'risk_score_imp_beta',
                      y = 'mean_meth',
                      xlab = "Risk score",
                      ylab = "Mean methylation",
                      model = 'seperate',
                      group = "CurrentPTSD", 
                      legend = '',
                      title = "Covariate adjusted and scaled\n with beta",
                      fig_legend = FALSE
                      )
# p_meth_sep


# p_meth_sep_l <- lm_plot(df = adj_df, x = 'risk_score',
#                       y = 'mean_meth',
#                       xlab = "Risk score",
#                       ylab = "Mean methylation",
#                       model = 'seperate',
#                       group = "LifetimePTSD",
#                       legend  = "Lifetime PTSD",
#                       title = "Covariate unadjusted")
# p_meth_sep_l



p_meth_adj_sep <- lm_plot(df = adj_sc_df, x = 'risk_score',
                          y = 'mean_meth',
                          xlab = "Risk score",
                          ylab = "",
                          model = 'seperate',
                          group = "CurrentPTSD",
                          legend = 'Current PTSD',
                          title = "Covariate adjusted and scaled")

# p_meth_adj_sep


# p_meth_adj_sep_l <- lm_plot(df = adj_df, x = 'risk_score',
#                           y = 'mean_meth',
#                           xlab = "Risk score",
#                           ylab = "",
#                           model = 'seperate',
#                           group = "LifetimePTSD",
#                           legend  = "Lifetime PTSD",
#                           title = "Covariate adjusted")

# p_meth_adj_sep_l
plot_grid(p_meth_sep, p_meth_adj_sep,
           labels = 'AUTO'
          )

# plot_grid(p_meth_sep, p_meth_adj_sep, 
#           p_meth_sep_l, p_meth_adj_sep_l, labels = 'AUTO'
#           )

```


```{r, fig.width=11, fig.height=4}
# # lm for risk score vs cumulative trauma
# case_model_ct <- lm(risk_score~Traumanum, data = cases)
# coef_case_ct <- coef(case_model_ct); coef_case_ct
# 
# control_model_ct <- lm(risk_score~Traumanum, data = controls)
# coef_control_ct <- coef(control_model_ct); coef_control_ct
# 
# 
# 
# lmr_ct <- lm(risk_score~Traumanum, data = adj_df)
# summary(lmr_ct)
# p_ct <- lm_plot(df = adj_df, x = 'Traumanum',
#                 y = 'risk_score', xlab = "Cumulative trauma",
#                 ylab = 'Risk score')
# # p_ct
# 
# 
# p_ct_sep <- lm_plot(df = adj_df, x = 'Traumanum',
#                 y = 'risk_score', xlab = "Cumulative trauma",
#                 ylab = 'Risk score', model = 'seperate')
# # p_ct_sep
# 
# cowplot::plot_grid(p_ct,NULL, p_ct_sep,
#                    nrow = 1,
#                    rel_widths = c(1,0.05, 1.5),
#                    labels = c("A", "", "B"))

```

# Cumulative trauma

```{r, fig.width=11, fig.height=4}

p_ct_meth_sep <- lm_plot(df = adj_df, x = 'TraumaNumber',
                         y = 'mean_meth', xlab = "Cumulative trauma",
                         ylab = 'Mean methylation',
                         model = 'seperate',
                         group = "CurrentPTSD",
                         legend = "Current PTSD",
                         title = "Covariate adjusted"
)
# p_ct_meth_sep


# p_ct_meth_sep_l <- lm_plot(df = unadj_df, x = 'TraumaNumber',
#                          y = 'mean_meth', xlab = "",
#                          ylab = 'Mean methylation',
#                          model = 'seperate',
#                          group = "LifetimePTSD",
#                          legend = "Lifetime PTSD",
#                          title = "Covariate unadjusted"
#                          )
# # p_ct_meth_sep_l

p_ct_meth_sep_adj <- lm_plot(df = adj_sc_df, x = 'TraumaNumber',
                         y = 'mean_meth', xlab = "Cumulative trauma",
                         ylab = '',
                         model = 'seperate',
                         group = "CurrentPTSD",
                         legend = "Current PTSD",
                         title = "Covariate adjusted and scaled"
)
# p_ct_meth_sep_adj


# p_ct_meth_sep_adj_l <- lm_plot(df = adj_df, x = 'TraumaNumber',
#                            y = 'mean_meth', xlab = "Cumulative trauma",
#                            ylab = '',
#                            model = 'seperate',
#                            group = "LifetimePTSD",
#                            legend = "Lifetime PTSD",
#                            title = "Covariate adjusted"
# )
# # p_ct_meth_sep_adj_l

plot_grid(p_ct_meth_sep, p_ct_meth_sep_adj,
           labels = 'AUTO'
          )
# plot_grid(p_ct_meth_sep, p_ct_meth_sep_adj, 
#           p_ct_meth_sep_l, p_ct_meth_sep_adj_l, labels = 'AUTO'
#           )

```


```{r, fig.width=11, fig.height=4}
# childhood trauma and mean methylation
lmr_cht_meth <- lm(mean_meth~Childhood_Mt, data = adj_df)
summary(lmr_cht_meth)

p_cht_meth <- lm_plot(df = adj_df, x = 'Childhood_Mt',
                y = 'mean_meth', xlab = "Childhood maltreatment",
                ylab = 'Mean methylation')
# p_cht_meth


p_cht_meth_sep <- lm_plot(df = adj_df, x = 'Childhood_Mt',
                y = 'mean_meth', xlab = "Childhood maltreatment",
                ylab = '', model = 'seperate')
# p_cht_meth_sep

cowplot::plot_grid(p_cht_meth,NULL, p_cht_meth_sep,
                   nrow = 1,
                   rel_widths = c(1,0.05, 1.5),
                   labels = c("A", "", "B"))

```


```{r}

```


### Density plot function

```{r}
make_densityplot <- function(df, x, y, xlab, ylab, fig_legend = TRUE,
                             title = ''){
  
  p <- ggplot(data=df, aes_string(x=x, group=y, fill=y)) +
    geom_density(adjust=1.5, alpha=.8) +
    theme_classic()
  
  if(fig_legend){
    p <- p + scale_fill_manual("Current PTSD",
                      values  = c("#0072B5FF", "#BC3C29FF"))
  }else{
    p <- p + scale_fill_manual("Current PTSD",
                      values  = c("#0072B5FF", "#BC3C29FF")) +
      theme(legend.position = 'none')
  }

  p <- p + ggtitle(title) +
    xlab(xlab) +
    ylab(ylab)+ th
}
```


```{r, fig.width=10, fig.height=5}

# for risk scores
p1_d <- make_densityplot(df = adj_sc_df, x = 'risk_score_imp_beta',
                         y = 'CurrentPTSD',
                         xlab = 'Risk score',
                         ylab = "Density",
                         fig_legend = FALSE,
                         title = "Covariate adjusted")
# p1_d

# for risk scores
p1_sc_d <- make_densityplot(df = adj_sc_df, x = 'risk_score',
                         y = 'CurrentPTSD',
                         xlab = 'Risk score',
                         ylab = "",
                         title = "Covariate adjusted and scaled")
# p1_sc_d



# for mean methylation
# p2_d <- make_densityplot(df = adj_df, x = 'mean_meth',
#                          y = 'CurrentPTSD',
#                          xlab = 'Mean metylation',
#                          ylab = "Density",
#                          fig_legend = FALSE)
# p2_d


cowplot::plot_grid(p1_d, p1_sc_d, rel_widths = c(1.3, 2),
                   labels = 'AUTO')



```


## Box plot function
```{r}
make_boxplot <- function(df, x, y, xlab, ylab,
                         test_pval, y_pos, title = ""){
  p <- df %>%
    ggplot(aes_string(x = x, y = y, fill = x)) +
    geom_boxplot(lwd = 1) +
    scale_fill_manual(values = c("#0072B5FF", "#BC3C29FF")) +
    geom_jitter(size=1.7, alpha=0.9) +
    theme_classic()+
    theme(legend.position="none") + th +

    ggtitle(title) +
    xlab(xlab) +
    ylab(ylab)

  # add significance
  p + geom_signif(
    annotation = formatC(test_pval, digits = 2),
    y_position = y_pos , xmin = 1, xmax = 2, size = 1,
    textsize = 5, tip_length = c(0.15, 0.02))
}

```


# ---------------------------------------------
```{r, fig.width=12, fig.height=4}
# wilcoxon test for difference is risk score between groups
wcox_test <- wilcox.test(as.numeric(adj_sc_df$CurrentPTSD),
                         adj_sc_df$risk_score_imp_beta)
wcox_test
wcox_test$p.value

p1_box <- make_boxplot(df = adj_sc_df, x = 'CurrentPTSD',
                       y = 'risk_score_imp_beta',
                       xlab = 'Current PTSD',
                       ylab = 'Risk score',
                       test_pval = wcox_test$p.value,
                       y_pos = 5, 
                       title = "Covariate adjusted"
                       )
p1_box


wcox_test_sc <- wilcox.test(as.numeric(adj_sc_df$CurrentPTSD),
                         adj_sc_df$risk_score)
wcox_test_sc
wcox_test_sc$p.value

p1_box_sc <- make_boxplot(df = adj_sc_df, x = 'CurrentPTSD',
                       y = 'risk_score',
                       xlab = 'Current PTSD',
                       ylab = 'Risk score',
                       test_pval = wcox_test_sc$p.value,
                       y_pos = .54,
                       title = "Covariate adjusted \nand scaled"
                       )
p1_box_sc


# --------------------------
wcox_test_meth <- wilcox.test(as.numeric(adj_df$CurrentPTSD),
                              adj_df$mean_meth)
wcox_test_meth
wcox_test_meth$p.value
# Plot risk scores for ptsd


p2_box <- make_boxplot(df = adj_df, x = 'CurrentPTSD',
                       y = 'mean_meth',
                       xlab = 'Current PTSD',
                       ylab = 'Mean methylation',
                       test_pval = wcox_test_meth$p.value,
                       y_pos  = .54,
                       title = "Covariate adjusted"
                       )
p2_box

plot_grid(p1_box, p1_box_sc, p2_box, nrow = 1)

```



```{r}
# get cpgs

cpgs_df <- adj_df %>%
  select(starts_with(c('cg','ch.')))
dim(cpgs_df)
head(cpgs_df[, 1:5])

cpgs <- names(cpgs_df)


# Check number of genes that significant adj_df are annotated to
check <- getMappedEntrezIDs(sig.cpg = cpgs)

length(check$sig.eg)


library(IlluminaHumanMethylationEPICanno.ilm10b2.hg19)
gst <- gometh(sig.cpg=cpgs, collection="GO",
              plot.bias=TRUE)

topGSA(gst)

```


